{ \fontfamily{ptm}\selectfont First Order Nonsmooth Optimization: Unifying Nesterov's acceleration and its applications }

### 2025-01-27 Notebook LaTEX/Proposal/presentation.tex
>w<, fixing cringe shit
>w<, changes and fixes

                \item It finds application in machine learning and it accelerates the convergence of Variance Reduced Method (A type of incremental method that is not slower than the exact counter part). 

@@

                \item see (3.1, 4.1) in Lee et al. \cite{lee_geometric_2021} and Ahn and Sra \cite{ahn_understanding_2022} for graphical visualization of similar triangle form. 

@@

        \begin{frame}{$\mu$ estimation graph by FR-WAPG}

@@

                \item $M, N$ are constants. $A \in \RR^{M\times N}$ ahs i.i.d random entries from a standard normal distribution. 

@@

                \item $x_0\in \RR^N$ is the initial guess with i.i.d randomed variable from a standard normal distributions. Same initial guess shared by FR-WAPG, V-FISTA, M-FISTA. 

@@

            There is one detail that our R-WAPG doesn't incorperate on all Euclidean variants of FISTA. 

@@

                \item Algorithm 5, Chambolle, Pock \cite{chambolle_introduction_2016} captures several variants of FISTA and it assumes that $F = f + g$ where $f, g$ has strong convexity constant $\mu_f \ge 0, \mu_g \ge 0$ respective so $F$ is $\mu := \mu_f + \mu_g \ge 0$ strongly convex. 

@@

            Fast linear convergence is possible if any one of $f, g$ of the function is strongly convex but this is not yet a prediction of R-WAPG. 

@@

@@

@@

@@

            \begin{proposition}[Inner loop, context but not strongly convex]\label{prop:inner-loop-complexity-cnvx}

@@

                    \item the initial guess for $\mathbb M$ is $z_{0, k} = x_{k - 1}$, 
                    \item $F$ has bounded level set.

@@

        \begin{frame}{Aggregated complexity}

@@

@@

@@

        \begin{frame}{Outerloop complexity under relative errors}
            \begin{theorem}[Outer loop complexity under criterion C2]\label{theorem:catalyst2-outer-loop}

@@

            Observe that it does't require knowledge about $F^*$. 

@@

    \subsection{Direction of future works}

>>>>

                \item It finds application in machine learning, and it accelerates the convergence of Variance Reduced Method (A type of incremental method that is not slower than the exact counterpart). 

@@

                \item See (3.1, 4.1) in Lee et al. \cite{lee_geometric_2021} and Ahn and Sra \cite{ahn_understanding_2022} for graphical visualization of similar triangle form. 

@@

        \begin{frame}{FR-WAPG $\mu$ estimation graph}

@@

                \item $M, N$ are constants. $A \in \RR^{M\times N}$ has i.i.d random entries from a standard normal distribution. 

@@

                \item $x_0\in \RR^N$ is the initial guess with i.i.d random variable from a standard normal distribution. Same initial guess shared by FR-WAPG, V-FISTA, M-FISTA. 

@@

            There is one detail that our R-WAPG doesn't incorporate on all Euclidean variants of FISTA. 

@@

                \item Algorithm 5, Chambolle, Pock \cite{chambolle_introduction_2016} captures several variants of FISTA, and it assumes that $F = f + g$ where $f, g$ has strong convexity constant $\mu_f \ge 0, \mu_g \ge 0$ respective so $F$ is $\mu := \mu_f + \mu_g \ge 0$ strongly convex. 

@@

            Fast linear convergence is possible if any one of $f, g$ of the function is strongly convex, but this is not yet a prediction of R-WAPG. 

@@

            Ths is Theorem 3.1 in Lin et al. \cite{lin_universal_2015}. 

@@

            This is Theorem 3.3 from Lin et al. \cite{lin_universal_2015}. 

@@

            The followings are Proposition 3.2, 3.4 from Lin et al. \cite{lin_universal_2015}. 

@@

            \begin{proposition}[Inner loop, convex but not strongly convex]\label{prop:inner-loop-complexity-cnvx}

@@

                    \item the initial guess for $\mathbb M$ is $z_{0, k} = x_{k - 1}$ and $F$ has bounded level set.

@@

        \begin{frame}{Total complexity}

@@

                Let the initial estimate be $x_0 \in \RR^n$, fix parameters $\kappa > 0$ and $\alpha_0 = \sqrt{q}$ where $q = \mu/(\kappa + \mu)$. 

@@

                        & \text{find } \alpha_k \in (0, 1) \text{ such that } \alpha_k^2 = (1 - \alpha_k)\alpha_{k - 1}^2 + q\alpha_k,

@@

        \begin{frame}{Outer loop complexity under relative errors}
            \begin{theorem}[Outer loop complexity under criterion C2]\label{thm:catalyst2-outer-comp}

@@

            Observe that it doesn't require knowledge about $F^*$. 

@@

        \begin{frame}{Inner loop warm start under criterion C2}
            \begin{assumption}[Warm start conditions for the inner loop using C2]\label{ass:warm-start-c2}
                For minimization sub problem $\mathcal M^{\kappa^{-1}}_F(z; y_k)$ solved by $\mathbb M$, we use the following warm start condition for $z_{k, 0}$ under termination criterion C2 (Definition \ref{def:catalyst-termination-c2}): 
                    \item If $F$ is smooth then choose $z_k = y_k$. 
                    \item Else it's composite, define $f_k(x) = f(x) + \frac{\kappa}{2}\Vert x - y_k\Vert^2$ and do 
                        z_{k, 0} = \hprox_{\eta g}\left(
                            y_k - \eta\nabla f_\kappa(y_k)
                        \right) \quad \text{with } \eta = 1/(\kappa + L). 
                See pg 13 of Lin et al. for more detail about warm starting strategies of $\mathbb M$ under Termination criterion C1. 
        \begin{frame}{Inner loop complexity under C2}
            \begin{theorem}[Inner loop complexity with relative error C2]\label{thm:inner-loop-comp-c2}
                Consider $\mathbb M$ under the settings of Assumption \ref{ass:linear-convergence-inner-loop} and Assumption \ref{ass:warm-start-c2}, let the relative error sequence $(\delta_k)_{k \ge0}$ be given by Theorem \ref{thm:catalyst2-outer-comp}, and let $\mathbb A$ be defined as in Definition \ref{def:catalyst-relative}. 
                Then the sequence $(z_{k, t})_{t \ge 0}$ generated by $\mathbb M$ has $T_{k + 1} = \min\left\lbrace t \ge 0 : z_{k, t} \in \widetilde{\mathcal J}_{\kappa^{-1}F}^\delta y_k\right\rbrace$ satisfies: 
                    T_{k + 1} &\le \tau_{\mathbb M}^{-1}
                    \log\left(
                        4 C_{\mathbb M}\frac{L + \kappa}{\kappa}\frac{2 - \sqrt{q}}{q}
                    \right) \quad \text{when}\quad \mu > 0, 
                    T_{k + 1} &\le \tau_{\mathbb M}^{-1}
                    \log\left(
                        4 C_{\mathbb M}\frac{L + \kappa}{\kappa}(k + 2)^2
                    \right) \quad \text{when}\quad \mu = 0.  
                This is Corollary 16 in Lin et al. \cite{lin_catalyst_2018} but phrased in our notations. 
        \begin{frame}{Total complexity under C2}
            The total number of iteration $N_{\mathbb M}$ experienced by $\mathbb M$ for $\mathbb A$ to produce $x_k$ such that $F(x_k) - F^* \le \epsilon$ when using termination criterion C2, warm start and error sequence strategies in Theorem \ref{thm:catalyst2-outer-comp} has: 
                \item $N_{\mathbb M} \le \widetilde{\mathcal O}\left(\tau_{\mathbb M}^{-1}\sqrt{\kappa/\epsilon}\log(\epsilon^{-1})\right)$ if $\mu = 0$, 
                \item $N_{\mathbb N}\le \widetilde{O}\left((\tau_{\mathbb M}\sqrt{q})^{-1}\log(\epsilon^{-1})\right)$ if $\mu > 0$. 
            In addition, termination criterion C2 has advantages: 
                \item $\delta_k$ doesn't requite $F^*$. 
                \item Doesn't require bounded level set assumption for inner loop complexity bound. 
                \item Generalize to non-convex settings which was also the focus of Paquette's paper on Catalyst \cite{paquette_catalyst_2018}. 
    \subsection{Future works for Catalyst Acceleration}

### 2025-01-27 Notebook LaTEX/Proposal/proposal.tex
>w<, changes and fixes

    \subsection{The second Catalyst Acceleration paper}\label{ssec:catalyst-second-paper}
        Lin et al.'s second paper on Catalyst Acceleration \cite{lin_catalyst_2018} describes the new option of relative termination criterion (which we stated here in Definition \ref{def:catalyst-termination-c2}) for $\mathbb M$ to evaluate the inexact solutions of $\mathcal J_{\kappa^{-1}F}^{\epsilon_k} y_k$. 
        The relative termination condition has advantages: it gives easier convergence analysis, it doesn't need knowledge of $F^*$, and it can generalize to the nonconvex case. 
        In addition, they also improved warm start strategies for $\mathbb M$ at the end of Section 3. 
        Both these ideas together provides a tighter and easier complexity analysis for the Catalyst Acceleration. 
        \par
        We summarize these innovations in Lin et al.'s second Catalyst paper \cite{lin_catalyst_2018} in this section. 
        The following definition defines relative termination criteria for $\mathbb M$. 
        \begin{definition}[Relative termination criterion C2]\label{def:catalyst-termination-c2}
            Take $F$ as given by Assumption \ref{ass:catalyst1}. 
            Given any $\delta \in (0, 1]$, $\kappa > 0$ and $x \in \RR^n$, the relative criterion C2 of the inexact resolvent is defined by: 
            \begin{align*}
                \widetilde{\mathcal J}_{\kappa^{-1}F}^\delta (x)
                := 
                \left\lbrace
                    z \in \RR^n \left| \;
                        \mathcal M_F^{\kappa^{-1}}(z; x) - 
                        \mathcal M^*_{F, \kappa^{-1}}(z; x) 
                        \le \frac{\kappa\delta}{2}\Vert x - z\Vert^2
                    \right.
                \right\rbrace. 
            \end{align*}
        \end{definition}
        \begin{remark}
            Observe that, if we set $\epsilon = \delta\kappa/2\Vert x - z\Vert^2$ then $\widetilde{\mathcal J}_{\kappa^{-1}F}^{\delta} (x) = \mathcal J_{\kappa^{-1}F}^\epsilon (x)$. 
            The relative inexact condition can be interpreted as an adaptive inexactness condition.     
            Observe that $\mathcal J_{\kappa^{-1}F} (x) \subseteq \widetilde{\mathcal J}_{\kappa^{-1}F}^{\delta}(x)$ hence the set is nonempty. 
        \end{remark}
        Stated by Lin, the following lemma provides the sufficient condition for evaluating inexact proximal point up to accuracy $\epsilon$. 
        \begin{lemma}[Sufficient condition for C1]\label{lemma:sufficient-c1}
            Consider smooth plus nonsmooth objective $F := f + g$ with $g:\RR^n \rightarrow \overline \RR, F:\RR^n \rightarrow \RR$ closed and convex, and $F$ is $\mu\ge 0$ strongly convex and $L$-Lipschitz smooth. 
            With arbitrary $x \in \RR^n$ fixed, the model function is additive composite of the form: 
            \begin{align*}
                \mathcal M_F^{\kappa^{-1}}(z; x) := 
                \underbrace{f(z) + \frac{\kappa}{2}\Vert z - x\Vert^2}_{=: f_\kappa (z)}
                 + g(z). 
            \end{align*}
            For any $z \in \RR^n$, define proximal gradient point 
            \begin{align*}
                \bar z = \hprox_{\eta g}(z - \nabla f_\kappa(z)) 
                \quad 
                \text{ with } \quad  \eta = 1/(\kappa + L). 
            \end{align*}
            Then it has 
            \begin{align*}
                \frac{1}{\eta}\Vert z - \bar z\Vert
                \le \sqrt{2\kappa \epsilon} \implies 
                \bar z \in \mathcal 
                J_{\kappa^{-1}F}^\epsilon (x).
            \end{align*}
        \end{lemma}
        \begin{remark}
            This is Lemma 2 in Lin et al. \cite{lin_catalyst_2018}. 
        \end{remark}

@@

                Consider $\mathbb M$ under the settings of Assumption \ref{ass:linear-convergence-inner-loop} and Assumption \ref{ass:warm-start-c2}, let the relative error sequence $(\delta_k)_{k \ge0}$ be given by Theomrem \ref{thm:catalyst2-outer-comp}, and let $\mathbb A$ be defined as in Definition \ref{def:catalyst-all-in-one}. 

@@

            The total number of iteration $N_{\mathbb M}$ experienced by $\mathbb M$ for $\mathbb A$ to produce $x_k$ such that $F(x_k) - F^* \le \epsilon$ when using termination criterion C2, warm start and error sequence strategies \ref{thm:catalyst2-outer-comp} has: 

>>>>

    \subsection{The second Catalyst Acceleration paper}\label{ssec:catalyst-second-paper}
        Lin et al.'s second paper on Catalyst Acceleration \cite{lin_catalyst_2018} describes the new option of relative termination criterion (which we stated here in Definition \ref{def:catalyst-termination-c2}) for $\mathbb M$ to evaluate the inexact solution of $\mathcal J_{\kappa^{-1}F}^{} y_k$. 
        The relative termination condition has advantages: it gives easier convergence analysis, it doesn't need knowledge of $F^*$, and it can generalize to the nonconvex case. 
        In addition, they also improved warm start strategies for $\mathbb M$ at the end of Section 3 for absolute termination criterion in Definition \ref{def:catalyst-termination-c1}.
        Both these ideas together provides a tighter and easier complexity analysis for the Catalyst Acceleration. 
        \par
        In this section, we summarize results regarding criterion C2 in Lin et al.'s second Catalyst paper \cite{lin_catalyst_2018}. 
        The following definition defines relative termination criteria for $\mathbb M$. 
        \begin{definition}[Relative termination criterion C2]\label{def:catalyst-termination-c2}
            Take $F$ as given by Assumption \ref{ass:catalyst1}. 
            Given any $\delta \in (0, 1]$, $\kappa > 0$ and $x \in \RR^n$, the relative criterion C2 of the inexact resolvent is defined by: 
            \begin{align*}
                \widetilde{\mathcal J}_{\kappa^{-1}F}^\delta (x)
                := 
                \left\lbrace
                    z \in \RR^n \left| \;
                        \mathcal M_F^{\kappa^{-1}}(z; x) - 
                        \mathcal M^*_{F, \kappa^{-1}}(z; x) 
                        \le \frac{\kappa\delta}{2}\Vert x - z\Vert^2
                    \right.
                \right\rbrace. 
            \end{align*}
        \end{definition}
        \begin{remark}
            Observe that, if we set $\epsilon = \delta\kappa/2\Vert x - z\Vert^2$ then $\widetilde{\mathcal J}_{\kappa^{-1}F}^{\delta} (x) = \mathcal J_{\kappa^{-1}F}^\epsilon (x)$. 
            The relative inexact condition can be interpreted as an adaptive inexactness condition.     
            Observe that $\mathcal J_{\kappa^{-1}F} (x) \subseteq \widetilde{\mathcal J}_{\kappa^{-1}F}^{\delta}(x)$ hence the set is nonempty. 
        \end{remark}
        Stated by Lin, the following lemma provides the sufficient condition for evaluating inexact proximal point up to accuracy $\epsilon$. 
        \begin{lemma}[Sufficient condition for C1]\label{lemma:sufficient-c1}
            Consider smooth plus nonsmooth objective $F := f + g$ with $g:\RR^n \rightarrow \overline \RR, F:\RR^n \rightarrow \RR$ closed and convex, and $F$ is $\mu\ge 0$ strongly convex and $L$-Lipschitz smooth. 
            With arbitrary $x \in \RR^n$ fixed, the model function is additive composite of the form: 
            \begin{align*}
                \mathcal M_F^{\kappa^{-1}}(z; x) := 
                \underbrace{f(z) + \frac{\kappa}{2}\Vert z - x\Vert^2}_{=: f_\kappa (z)}
                 + g(z). 
            \end{align*}
            For any $z \in \RR^n$, define proximal gradient point 
            \begin{align*}
                \bar z = \hprox_{\eta g}(z - \nabla f_\kappa(z)) 
                \quad 
                \text{ with } \quad  \eta = 1/(\kappa + L). 
            \end{align*}
            Then it has 
            \begin{align*}
                \frac{1}{\eta}\Vert z - \bar z\Vert
                \le \sqrt{2\kappa \epsilon} \implies 
                \bar z \in \mathcal 
                J_{\kappa^{-1}F}^\epsilon (x).
            \end{align*}
        \end{lemma}
        \begin{remark}
            This is Lemma 2 in Lin et al. \cite{lin_catalyst_2018}. 
        \end{remark}

@@

                Consider $\mathbb M$ under the settings of Assumption \ref{ass:linear-convergence-inner-loop} and Assumption \ref{ass:warm-start-c2}, let the relative error sequence $(\delta_k)_{k \ge0}$ be given by Theorem \ref{thm:catalyst2-outer-comp}, and let $\mathbb A$ be defined as in Definition \ref{def:catalyst-all-in-one}. 

@@

            The total number of iteration $N_{\mathbb M}$ experienced by $\mathbb M$ for $\mathbb A$ to produce $x_k$ such that $F(x_k) - F^* \le \epsilon$ when using termination criterion C2, warm start and error sequence strategies in Theorem \ref{thm:catalyst2-outer-comp} has: 
